---
title: "STA410 Homework 1"
subtitle: "student number: 1003942326"
author: "Yulin WANG"
date: "`r Sys.Date()`"
output: pdf_document
---
\newpage
# Question1

## (a)
Plug $\hat{Z}=A_mZA_n^T$ in:
$$
\begin{aligned}
D_m^{-1}A_m^T\hat{Z}A_nD_n^{-1} &= D_m^{-1}A_m^T\{A_mZA_n^T\}A_nD_n^{-1} \\
&=D_m^{-1}\{A_m^TA_m\}Z\{A_n^TA_n\}D_n^{-1}\\
&=\{D_m^{-1}D_m\}Z\{D_nD_n^{-1}\} \ \ \ \ \text{since}\ A_m^TA_m=D_m \ \text{and}\ A_n^TA_n=D_n \\ 
&=Z
\end{aligned}
$$


## (b)
### hard-thresholding function

```{r}
library('dtt')
#hard-thresholding function

hard_thres <- function(dctmat,lambda) {
  # if lambda is missing, set it to the 0.8 quantile of abs(dctmat)
  if(missing(lambda)) lambda <- quantile(abs(dct),0.8)
  a <- dctmat[1,1]
  dctmat1 <- ifelse(abs(dctmat)>lambda,dctmat,0)
  dctmat1[1,1] <- a
  # inverse DCT to obtain denoised image "clean"
  clean <- mvdct(dctmat1,inverted=T)
  clean <- ifelse(clean<0,0,clean)
  clean <- ifelse(clean>1,1,clean)
  clean
  }

```


### soft-thresholding function

```{r}
library('dtt')
#soft-thresholding function

soft_thres <- function(dctmat,lambda) {
  if(missing(lambda)) lambda <- quantile(abs(dct),0.8)
  a <- dctmat[1,1]
  dctmat1 <- sign(dctmat)*pmax(abs(dctmat)-lambda,0)
  dctmat1[1,1] <- a
  clean <- mvdct(dctmat1,inverted=T)
  clean <- ifelse(clean<0,0,clean)
  clean <- ifelse(clean>1,1,clean)
  clean
  }
```


\newpage
## (c)

### original image

```{r}
#load the 'boats.txt' and create the original image
boats <- matrix(scan("boats.txt"),ncol=256,byrow=T)
image(boats, axes=F, col=grey(seq(0,1,length=256)),sub="original image")
```

### create four denoised images by using hard-thresholding with $\lambda=5,10,20,40$

```{r}
#create the images by hard-thres
boat1 <- hard_thres(mvdct(boats),lambda=5)
boat2 <- hard_thres(mvdct(boats),lambda=10)
boat3 <- hard_thres(mvdct(boats),lambda=20)
boat4 <- hard_thres(mvdct(boats),lambda=40)
image1<-image(boat1, axes=F, col=grey(seq(0,1,length=256)),
              sub="hard-thresholding with lambda = 5",asp=1)
image2<-image(boat2,axes=F, col=grey(seq(0,1,length=256)),
              sub="hard-thresholding with lambda = 10",asp=1)
image3<-image(boat3,axes=F, col=grey(seq(0,1,length=256)),
              sub="hard-thresholding with lambda = 20",asp=1)
image4<-image(boat4,axes=F, col=grey(seq(0,1,length=256)),
              sub="hard-thresholding with lambda = 40",asp=1)
```

\newpage
### create four denoised images by using soft-thresholding with $\lambda=5,10,20,40$

```{r}
#create the images by soft-thres
boat5 <- soft_thres(mvdct(boats),lambda=5)
boat6 <- soft_thres(mvdct(boats),lambda=10)
boat7 <- soft_thres(mvdct(boats),lambda=20)
boat8 <- soft_thres(mvdct(boats),lambda=40)
image5<-image(boat5, axes=F, col=grey(seq(0,1,length=256)),
              sub="soft-thresholding with lambda = 5",asp=1)
image6<-image(boat6,axes=F, col=grey(seq(0,1,length=256)),
              sub="soft-thresholding with lambda = 10",asp=1)
image7<-image(boat7,axes=F, col=grey(seq(0,1,length=256)),
              sub="soft-thresholding with lambda = 20",asp=1)
image8<-image(boat8,axes=F, col=grey(seq(0,1,length=256)),
              sub="soft-thresholding with lambda = 40",asp=1)
```



### conclusion
Compared the above eight denoised images and the original one, we can find that for both hard-thresholding and soft-thresholding functions, the smaller $\lambda$, the better image denoised, and the hard-thresholding function works better than soft-thresholding for the same $\lambda$. 






\newpage
# Question2
## (a)
Since  $U$ and $V$ are independent, so 
$$g(s)=E(s^X)=E(s^{U+2V})=E(s^U \cdot s^{2V})=E(s^U) \cdot E(s^{2V})$$
And we have:
$$
\begin{aligned}
E(s^U) &= \sum_{k=0}^{\infty}s^kP(U=k) \\
&= \sum_{k=0}^{\infty}s^k\cdot\frac{e^{-\lambda_u} \lambda_u^k}{k!} \ \ \ \ \text{since} \  U\sim Pois(\lambda_u)\\
&= e^{-\lambda_u}\cdot\sum_{k=0}^{\infty}\frac{s^k\lambda_u^k}{k!} \\
&= e^{-\lambda_u}\cdot\sum_{k=0}^{\infty}\frac{(\lambda_u s)^k}{k!} \\
&= e^{-\lambda_u}\cdot e^{\lambda_u s} \\
&= e^{\lambda_u (s-1)}
\end{aligned}
$$
$$
\begin{aligned}
E(s^{2V}) &= \sum_{k=0}^{\infty}s^{2k}P(V=k) \\
&= \sum_{k=0}^{\infty}s^{2k}\cdot\frac{e^{-\lambda_v} \lambda_v^k}{k!} \ \ \ \ \text{since} \  V\sim Pois(\lambda_v) \\
&= e^{-\lambda_v}\cdot\sum_{k=0}^{\infty}\frac{s^{2k}\lambda_v^k}{k!} \\
&= e^{-\lambda_v}\cdot\sum_{k=0}^{\infty}\frac{(\lambda_v s^2)^k}{k!} \\
&= e^{-\lambda_v}\cdot e^{\lambda_v s^2} \\
&= e^{\lambda_v (s^2-1)}
\end{aligned}
$$
Thus,
$$
\begin{aligned}
g(s)=E(s^X)&=e^{\lambda_u (s-1)} \cdot e^{\lambda_v (s^2-1)}\\
&=\text{exp}[ \,\lambda_u (s-1)+\lambda_v (s^2-1)] \, 
\end{aligned}
$$

\newpage
## (b)
Using the inequality, for any $s>1$, we can define $M$ to satisfy
$$\frac{\text{exp}[ \,\lambda_u (s-1)+\lambda_v (s^2-1)] \,}{s^M} = \epsilon$$
then we have
$$M(s)=\frac{\lambda_u (s-1)+\lambda_v (s^2-1)-ln(\epsilon)}{ln(s)}$$
where $P(X\ge M(s))\le \epsilon$. Since this is true for any $s>1$, we can take $M(s)$ to be as small as possible:
$$M=\inf_{s>1}\frac{\lambda_u (s-1)+\lambda_v (s^2-1)-ln(\epsilon)}{ln(s)}$$
with $P(X\ge M)\le \epsilon$.

\newpage
## (c)

### The R function named "find_dist" that implements the algorithm.

```{r}
find_dist <- function(lambda_u,lambda_v,epsilon){ 
  # three inputs: lambda_u, lambda_v, and epsilon
  S <- c(1001:10000)/1000 #a discrete set of points S
  M <- min((lambda_u*(S-1)+lambda_v*(S^2-1)-log(epsilon))/log(S)) #value of M
  s <- exp(-2*pi*1i*c(0:(M-1))/M) #value of s
  gs <- exp(lambda_u*(s-1) + lambda_v*(s^2-1)) #g(s): pgf of X
  px <- Re(fft(gs,inverse=T))/M #evaluate p(X=x) by computing the inverse fft of g(s)
  r <- list(x=c(0:(M-1)),probs=px) 
  plot(r$x,r$probs,type="h",lwd=5,main="distribution of X", xlab="x",ylab="probability")
}
```

### (i) When $\lambda_u=1 \ \text{and} \ \lambda_v=5$
```{r}
find_dist(lambda_u=1,lambda_v=5,epsilon=10^(-5))
```

\newpage
### (ii) When $\lambda_u=0.1 \ \text{and} \ \lambda_v=2$
```{r}
find_dist(lambda_u=0.1,lambda_v=2,epsilon=10^(-5))
```





